---
title: "Stat4911PracticeProject"
author: "Armaan Sodhi, Max Margolis, Nathan Snyder, Gaurav Law"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
#install.packages("nyfcflights13")
library(nycflights13)
library(ggplot2)
library(dplyr)
library(randomForest)
library(tree)
library(lubridate)
library(grid)
```


```{r}
airlines <- nycflights13::airlines
airports <- nycflights13::airports
flights <- nycflights13::flights
planes <- nycflights13::planes
weather <- nycflights13::weather
```


```{r}
top10AirportsDest <- flights %>%
  group_by(dest) %>%
  count %>%
  arrange(desc(n)) %>%
  head(10)
```


```{r}
ggplot(top10AirportsDest, aes(x = reorder(dest, -n), y = n, fill = dest)) +
    geom_bar(stat = "identity") +
    labs(title = "Top 10 Airports by Destination",
         x = "Destination Airport",
         y = "Count") +
    scale_fill_brewer(palette = "Set3") +
    theme_minimal()
```

This visualization shows the 3 airports that are listed as the origin of a flight the most. The height of each bar represents the number of times (count) that each airport is listed as an origin. EWR (Newark Liberty National Airport) seems to be the top origin airport of the 3, followed by John F Kennedy and then La Guardia.


```{r}
AirportOrigin <- flights %>%
  group_by(origin) %>%
  count %>%
  arrange(desc(n)) %>%
  head(10)
```



```{r}
ggplot(AirportOrigin, aes(x = reorder(origin, -n), y = n, fill = origin)) +
    geom_bar(stat = "identity") +
  labs(title = "Top Airports by Origin",
       x = "Origin Airport",
       y = "Count") +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal()

```


```{r}
flights %>%
  group_by(dest) %>%
  count() %>%
  summarise(frequency = n,
            origin) %>%
  arrange(-frequency)

flights <- flights %>%
  drop_na()

weather <- weather %>%
  drop_na()

flights$delay <- TRUE

for (i in 1:nrow(flights)) {
  if (flights$dep_delay[i] >= 15) {
    flights$delay[i] = TRUE}
  else if (flights$dep_delay[i] >= 0 && flights$dep_delay[i] < 15) {
    flights$delay[i] = FALSE
  } else {
    flights$delay[i] = "Early"
  }
}

flights %>% 
  ggplot(aes(x = delay, y = dep_delay)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 16, alpha = 0.6) +
  geom_jitter(width = 0.2, alpha = 0.6) +
  labs(title = "Boxplot of delays",
       x = "Delay?",
       y = "Frequency") +
  theme_classic()


flights %>%
  count(delay)
```


```{r plane sizes}

planes$size <- "Small"

for (i in 1:nrow(planes)) {
  if (planes$seats[i] <= 25) {
    planes$size[i] = "Small"
    } else if (planes$seats[i] > 25 && planes$seats[i] <= 180) {
    planes$size[i] = "Medium"
    } else if (planes$size[i] > 180) {
    planes$size[i] = "Large"
  }
}

planes %>% 
  ggplot(aes(x = size, y = engines)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 16, alpha = 0.6) +
  geom_jitter(width = 0.2, alpha = 0.6) +
  labs(title = "Boxplot of sizes",
       x = "Size of plane",
       y = "Frequency") +
  theme_classic()

planes %>%
  count(size)

```


```{r}
flights <- flights %>%
  left_join(airlines, by = c("carrier"))

delays <- flights %>%
  group_by(name) %>%
  summarise(min_dep_delay = min(dep_delay),
            avg_dep_delay = mean(dep_delay),
            max_dep_delay = max(dep_delay),
            min_arr_delay = min(arr_delay),
            avg_arr_delay = mean(arr_delay),
            max_arr_delay = max(arr_delay),
            overall = avg_arr_delay - avg_dep_delay)
```



To look at for weather relationship with delay:
  humid
  

  wind_dir
  wind_speed
  precip
  visib


# Merging datasets
```{r}

merged <- flights %>%
  left_join(weather, by = c("origin", "year", "month", "day", "hour", "time_hour"))

merged <- merged %>%
  left_join(airlines, by = c("carrier"))

merged <- merged %>%
  left_join(airports, by = c(dest = "faa"))

merged <- merged %>% drop_na() # drop NA values

# Planes have no significance in model

# planes <- planes[, !names(planes) %in% "speed"]

#merged <- merged %>% left_join(planes, by = c("tailnum"))

merged <- merged %>%
  mutate(delay = as.character(delay)) %>%
  mutate(delay = case_when(
    delay == "TRUE" ~ 1,
    delay == "Early" ~ -1,
    delay == "FALSE" ~ 0,
))
```


## Merged flights + planes
```{r}
fl_pl <- flights %>%
  left_join(planes, by = c("tailnum"))

fl_pl <- fl_pl[, !names(fl_pl) %in% "speed"]

colSums(is.na(fl_pl))
fl_pl <- fl_pl %>%
  drop_na()
nrow(fl_pl)
```


#Gets delayed/non-delayed flights
```{r}
delayFlights <- merged %>%
  filter(merged$delay == "TRUE")

nonDelayFlights <- merged %>%
  filter(merged$delay == "FALSE")
```


#EDA with Delayed/non-delayed Flights
```{r}
mean(delayFlights$visib)
mean(nonDelayFlights$visib)
```


```{r}
heatmap_data <- flights %>%
  filter(!is.na(dep_delay), !is.na(hour)) %>% # Filter out missing delays and hours
  group_by(carrier, hour) %>% # Group by carrier and hour of day
  summarize(avg_delay = mean(dep_delay, na.rm = TRUE)) %>% # Calculate average delay
  ungroup()

heatmap_data <- heatmap_data %>%
  left_join(airlines, by = c("carrier" = "carrier")) # Join with 'airlines' for names

ggplot(heatmap_data, aes(x = hour, y = name, fill = avg_delay)) +
  geom_tile() +
  scale_fill_gradient(low = "blue", high = "red", na.value = "grey50", name = "Avg Delay (min)") +
  labs(
    title = "Heatmap of Average Flight Delays",
    x = "Hour of Day",
    y = "Airline",
    caption = "Data source: nycflights13"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


This heatmap shows the average delay for each airline throughout the day (by hour). The scale goes from blue-red with the more red meaning the greater the average delay. he average delay across all airlines seems to increase as the hour increases. Mesa Airlines appears to have a random spike at hour 10 where the average delay is much greater than the other hours for that same airline.


```{r}
weather_by_airport <- weather %>%
  filter(!is.na(temp), !is.na(dewp), !is.na(humid), !is.na(wind_dir),
         !is.na(wind_speed), !is.na(wind_gust), !is.na(precip),
         !is.na(pressure), !is.na(visib)) %>%
  group_by(origin, month) %>% # Group by airport and month
  summarize(
    avg_temp = mean(temp, na.rm = TRUE),
    avg_dewp = mean(dewp, na.rm = TRUE),
    avg_humid = mean(humid, na.rm = TRUE),
    avg_wind_dir = mean(wind_dir, na.rm = TRUE),
    avg_wind_speed = mean(wind_speed, na.rm = TRUE),
    avg_wind_gust = mean(wind_gust, na.rm = TRUE),
    avg_precip = mean(precip, na.rm = TRUE),
    avg_pressure = mean(pressure, na.rm = TRUE),
    avg_visib = mean(visib, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = starts_with("avg_"), names_to = "variable", values_to = "value") %>%
  mutate(variable = gsub("avg_", "", variable))

ggplot(weather_by_airport, aes(x = month, y = variable, fill = value)) +
  geom_tile() +
  scale_fill_gradient(low = "blue", high = "red", na.value = "grey50", name = "Average Value") +
  labs(
    title = "Heatmap of Weather Variables by Airport Over Time (by Month)",
    x = "Month",
    y = "Weather Variable",
    caption = "Data source: nycflights13"
  ) +
  facet_wrap(~origin) + # Facet by airport
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 10)
  )
```
This heatmap shows the average value each weather variable across 12 months for each of the three origin airports. (to be continued)





```{r}
size_counts <- planes %>%
  group_by(size) %>%
  summarize(count = n())

ggplot(size_counts, aes(x = size, y = count, fill = size)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Counts of Small, Medium, and Large Planes",
    x = "Plane Size",
    y = "Count",
    fill = "Plane Size"
  ) +
  theme_minimal()
```

This bar graph compares the number of small planes, medium planes, and large planes. The height of each bar is the number (or count) of each type of plane.Medium planes have by far the highest count, with small planes being the lowest.


```{r}

flights_with_size <- flights %>%
  inner_join(planes, by = "tailnum") 

size_by_airport_counts <- flights_with_size %>%
  group_by(origin, size) %>% # Group by airport (origin) and size
  summarize(count = n()) %>%
  ungroup()

ggplot(size_by_airport_counts, aes(x = origin, y = count, fill = size)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Counts of Small, Medium, and Large Planes by Airport",
    x = "Airport",
    y = "Count",
    fill = "Plane Size"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


This visualization shows the number of small, medium, and large planes for each of the 3 airports. The height of each bar is the number (or count) for each type of plane. EWR and LGA have more medium planes, while JFK has more large planes. All 3 airports have a lower amount of small planes.




```{r}
flights %>%
  count(delay) %>%
  ggplot(aes(x = delay, y = n, fill = delay)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Proportion of Flights by Delay Status",
    x = "Delay Status",
    y = "Number of Flights",
    fill = "Delay Status"
  ) +
  theme_minimal()
```


In this visualization, there are 3 different delay statuses of flights - True, False, or Early. This visualization shows the proportion of each of these statuses represented by bars, where the y-axis is the proportion. It appears that there are a much higher proportion of delayed flights than on time or early flights

## Delay By Airport


```{r}
flights %>%
  filter(delay == TRUE) %>%
  count(origin) %>%
  ggplot(aes(x = origin, y = n, fill = origin)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Number of Delayed Flights by Airport",
    x = "Airport",
    y = "Number of Delays",
    fill = "Airport"
  ) +
  theme_minimal()
```


In this visualization, there are 3 different delay statuses of flights - True, False, or Early. This visualization shows the number of each of these statuses represented by bars, where the y-axis is the count. It appears that there are a much higher proportion of delayed flights than on time or early flights.




## Delays by Carrier


```{r}
flights %>%
  filter(delay == TRUE) %>%
  count(carrier) %>%
  left_join(airlines, by = "carrier") %>%
  ggplot(aes(x = name, y = n, fill = name)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Number of Delayed Flights by Carrier",
    x = "Airline",
    y = "Number of Delays",
    fill = "Airline"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


This bar graph lists out each airline and shows the number of delays of each. Each airline corresponds to a bar and the height of each bar is the number of delays for that airline. Delta,ExpressJet,JetBlue, and United Air Lines seem to have the most delays.


## Delays Over Time



```{r}
flights %>%
  filter(delay == TRUE) %>%
  group_by(month) %>%
  summarize(count = n()) %>%
  ggplot(aes(x = month, y = count)) +
  geom_line(color = "red") +
  geom_point() +
  labs(
    title = "Number of Delays Over Time (by Month)",
    x = "Month",
    y = "Number of Delays"
  ) +
  theme_minimal()
```


This visualization shows the number of delays over time by month through 12 months. Each month has a dot for which the height represents the number of delays. These dots are connected by a red line. It seems the number of delays picks up at around month 3 and stays up there for the rest of the months.


## Delays because of Precipitation


```{r}
flights_with_weather <- flights %>%
  inner_join(weather, by = c("origin", "time_hour"))

flights_with_weather %>%
  filter(!is.na(delay)) %>%
  ggplot(aes(x = delay, y = precip, fill = delay)) +
  geom_boxplot() +
  labs(
    title = "Effect of Precipitation on Flight Delays",
    x = "Delay Status",
    y = "Precipitation (inches)",
    fill = "Delay Status"
  ) +
  theme_minimal()
```


## Delays by Time of Day


```{r}
flights %>%
  filter(delay == TRUE) %>%
  count(hour) %>%
  ggplot(aes(x = hour, y = n)) +
  geom_line(color = "blue") +
  geom_point() +
  labs(
    title = "Delays by Hour of the Day",
    x = "Hour",
    y = "Number of Delays"
  ) +
  theme_minimal()
```


This visualization shows the number of delays by hour of the day. Each hour has a dot for which the height is the number of delays. These dots are connected by the blue line. It seems that the beginning and ends of the day have less delays than the middle of the day, although there is a little dip at around hour 10.

## Delays by Plane Size


```{r}
flights_with_planes <- flights %>%
  inner_join(planes, by = "tailnum")

flights_with_planes %>%
  filter(!is.na(size)) %>%
  count(size, delay) %>%
  ggplot(aes(x = size, y = n, fill = delay)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Delays by Plane Size",
    x = "Plane Size",
    y = "Count",
    fill = "Delay Status"
  ) +
  theme_minimal()
```


This bar chart shows the delay status for each of the 3 sizes of planes.It uses the count of each represented by the height of the bars. "True" (or delayed) seems to be the heighest delay status for each of the 3 types.


## Heatmap of Delays (by Weather and Time)


```{r}
flights_with_weather %>%
  filter(delay == TRUE) %>%
  group_by(month.x, hour.x) %>%
  summarize(avg_precip = mean(precip, na.rm = TRUE)) %>%
  ggplot(aes(x = hour.x, y = month.x, fill = avg_precip)) +
  geom_tile() +
  scale_fill_gradient(low = "blue", high = "red") +
  labs(
    title = "Heatmap of Delays by Precipitation and Time",
    x = "Hour of Day",
    y = "Month",
    fill = "Average Precipitation"
  ) +
  theme_minimal()
```


## Barchart of Average Wind Speed by Delay Category


```{r}
ggplot(flights_with_weather, aes(x = delay, y = wind_speed, fill = delay)) + 
  stat_summary(fun = "mean", geom = "bar") + 
  labs(title = "Average Wind Speed by Flight Delay Category", 
       x = "Flight Delay Category", 
       y = "Average Wind Speed") +
  theme_minimal()

```


This barchart shows the average wind speed (at the origin airport) for each delay category, with the height of each bar representing the average wind speed. Delayed flights seem to have a slightly higher average wind speed than on time and early flights.


## Barchart of Average Temperature by Delay Category


```{r}
ggplot(flights_with_weather, aes(x = delay, y = temp, fill = delay)) + 
  stat_summary(fun = "mean", geom = "bar") + 
  labs(title = "Average Temperature by Flight Delay Category", 
       x = "Flight Delay Category", 
       y = "Average Temperature") +
  theme_minimal()
```


This barchart shows the average temperature (at the origin airport) for each delay category, with the height of each bar representing the average temperature. Delayed flights seem to have a higher average temperature than on time and early flights.


## Barchart of Average Humidity by Delay Category


```{r}
ggplot(flights_with_weather, aes(x = delay, y = humid, fill = delay)) + 
  stat_summary(fun = "mean", geom = "bar") + 
  labs(title = "Average Humidity by Flight Delay Category", 
       x = "Flight Delay Category", 
       y = "Average Humidity") +
  theme_minimal()
```


This barchart shows the average humidity (at the origin airport) for each delay category, with the height of each bar representing the average humidity. There does not appear to be much of a correlation here, although early flights seem to have slightly higher average humidity.


## Barchart of Differnce in Arrival Delay and Departure Delay by Airline


```{r}
delays %>% 
  ggplot(aes(x = name, y = overall)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Arrival Delay - Departure Delay by Airline", 
       x = "Airline", 
       y = "Delay Difference (Arrival - Departure)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```


This visualiztion shows the difference between the average arrival time and averague departure time for each airline. Alaska Airlines seems to be the most negative (meaning the arrival delay is much lower than the departure delay), while Frontier Airlines seems to be the most positive (meaning the arrival delay is greater than the departure delay)


## Barchart of Average Departure and Arrival Delay by Airline


```{r}
delays_long <- delays %>%
  gather(key = "delay_type", value = "avg_delay", avg_dep_delay, avg_arr_delay)

# Create the bar plot
ggplot(delays_long, aes(x = name, y = avg_delay, fill = delay_type)) +
  geom_bar(stat = "identity", position = "dodge") +  # position "dodge" puts bars side by side
  labs(title = "Average Departure and Arrival Delays by Airline",
       x = "Airline", y = "Average Delay (minutes)") +
  scale_fill_manual(values = c("skyblue", "orange")) +  # Customize bar colors
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
```


This visualization shows both the average departure delay and arrival delay for each airline. Frontier airlines appears to have a higher arrival and depature delay than the others. Alaskan and Hawaiian airlines both have a negative average arrival delay, meaning the arrive early on average.

#PNG images
```{r}
library(png)
url <- "https://logos-world.net/wp-content/uploads/2021/02/Alaska-Airlines-Logo-700x394.png"
filepath <- "alaskaAir.png"
download.file(url, filepath, mode = "wb")
alaskaAir <- readPNG(filepath)

url1 <- "https://logos-world.net/wp-content/uploads/2021/08/Hawaiian-Airlines-Logo-700x394.png"
filepath1 <- "hawaiianAir.png"
download.file(url1, filepath1, mode = "wb")
hawaiianAir <- readPNG(filepath1)

url2 <- "https://logos-world.net/wp-content/uploads/2023/01/Virgin-Atlantic-Logo-500x281.png"
filepath2 <- "Virgin.png"
download.file(url2, filepath2, mode = "wb")
virginAir <- readPNG(filepath2)

url3 <- "https://logos-world.net/wp-content/uploads/2023/05/Endeavor-Air-Logo-500x281.png"
filepath3 <- "Endeavor.png"
download.file(url3, filepath3, mode = "wb")
endeavorAir <- readPNG(filepath3)

url4 <- "https://1000logos.net/wp-content/uploads/2017/06/United-Airlines-Logo.png"
filepath4 <- "United.png"
download.file(url4, filepath4, mode = "wb")
unitedAir <- readPNG(filepath4)

url5 <- "https://logos-world.net/wp-content/uploads/2020/11/American-Airlines-Emblem.png"
filepath5 <- "American.png"
download.file(url5, filepath5, mode = "wb")
americanAir <- readPNG(filepath5)

url6 <- "https://logos-world.net/wp-content/uploads/2020/10/Southwest-Airlines-Emblem.png"
filepath6 <- "Southwest.png"
download.file(url6, filepath6, mode = "wb")
southwestAir <- readPNG(filepath6)

url7 <- "https://logos-world.net/wp-content/uploads/2021/08/Delta-Emblem.png"
filepath7 <- "Delta.png"
download.file(url7, filepath7, mode = "wb")
deltaAir <- readPNG(filepath7)

url8 <- "https://upload.wikimedia.org/wikipedia/commons/4/47/ExpressJet_Airlines%2C_LLC_-_Copy.png"
filepath8 <- "Express.png"
download.file(url8, filepath8, mode = "wb")
expressAir <- readPNG(filepath8)

url9 <- "https://1000logos.net/wp-content/uploads/2019/12/JetBlue-Airways-Logo.png"
filepath9 <- "JetBlue.png"
download.file(url9, filepath9, mode = "wb")
jetblueAir <- readPNG(filepath9)

url10 <- "https://www.aviatorcapital.com/wp-content/uploads/2017/11/Mesa-Airlines.png"
filepath10 <- "Mesa.png"
download.file(url10, filepath10, mode = "wb")
mesaAir <- readPNG(filepath10)

url11 <- "https://upload.wikimedia.org/wikipedia/commons/thumb/a/a6/US_Airways_Logo_2011.svg/800px-US_Airways_Logo_2011.svg.png"
filepath11 <- "USAirways.png"
download.file(url11, filepath11, mode = "wb")
usAirwaysAir <- readPNG(filepath11)

url12 <- "https://1000logos.net/wp-content/uploads/2023/06/SkyWest-Logo-500x281.png"
filepath12 <- "Skywest.png"
download.file(url12, filepath12, mode = "wb")
skywestAir <- readPNG(filepath12)

url13 <- "https://www.envoyair.com/wp-content/uploads/2024/03/Envoy-logo_blue-01-1.png"
filepath13 <- "Envoy.png"
download.file(url13, filepath13, mode = "wb")
envoyAir <- readPNG(filepath13)

url14 <- "https://logos-world.net/wp-content/uploads/2023/01/AirTran-Airways-Logo-500x281.png"
filepath14 <- "AirTran.png"
download.file(url14, filepath14, mode = "wb")
airtranAir <- readPNG(filepath14)

url15 <- "https://logos-world.net/wp-content/uploads/2021/03/Frontier-Airlines-Emblem.png"
filepath15 <- "Frontier.png"
download.file(url15, filepath15, mode = "wb")
frontierAir <- readPNG(filepath15)
```


```{r}
delays$name <- c(
  "AirTran",
  "Alaska",
  "American",
  "Delta",
  "Endeavor",
  "Envoy",
  "ExpressJet",
  "Frontier",
  "Hawaiian",
  "JetBlue",
  "Mesa",
  "SkyWest",
  "Southwest",
  "United",
  "USAirways",
  "Virgin"
)

library(grid)

logos <- list(
  Alaska = rasterGrob(alaskaAir, interpolate = TRUE),
  Hawaiian = rasterGrob(hawaiianAir, interpolate = TRUE),
  Virgin = rasterGrob(virginAir, interpolate = TRUE),
  Endeavor = rasterGrob(endeavorAir, interpolate = TRUE),
  United = rasterGrob(unitedAir, interpolate = TRUE),
  American = rasterGrob(americanAir, interpolate = TRUE),
  Southwest = rasterGrob(southwestAir, interpolate = TRUE),
  Delta = rasterGrob(deltaAir, interpolate = TRUE),
  ExpressJet = rasterGrob(expressAir, interpolate = TRUE),
  JetBlue = rasterGrob(jetblueAir, interpolate = TRUE),
  Mesa = rasterGrob(mesaAir, interpolate = TRUE),
  USAirways = rasterGrob(usAirwaysAir, interpolate = TRUE),
  SkyWest = rasterGrob(skywestAir, interpolate = TRUE),
  Envoy = rasterGrob(envoyAir, interpolate = TRUE),
  AirTran = rasterGrob(airtranAir, interpolate = TRUE),
  Frontier = rasterGrob(frontierAir, interpolate = TRUE)
)

logoGraph <- ggplot(delays, aes(x = avg_dep_delay, y = avg_arr_delay)) +
  geom_point(color = "white") + 
  xlim(0, max(delays$avg_dep_delay) + 2) +
  ylim(0, max(delays$avg_arr_delay) + 2) +
  labs(title = "Airline Delays", x = "Average Departure Delay", y = "Average Arrival Delay") +
  theme_minimal()

for (i in 1:nrow(delays)) {
  logoGraph <- logoGraph + annotation_custom(
    logos[[delays$name[i]]],
    xmin = delays$avg_dep_delay[i] - 1.5,
    xmax = delays$avg_dep_delay[i] + 1.5,
    ymin = delays$avg_arr_delay[i] - 1.5,
    ymax = delays$avg_arr_delay[i] + 1.5
  )
}

print(logoGraph)
```



```{r}
modelWeather <- glm(delay ~ temp + dewp + humid + wind_dir + wind_speed + wind_gust + precip + pressure + visib, data = merged)
summary(modelWeather)
```


We find that precip has a p value of 0.667210 which is greater than our alpha value of .05 so we would say that precipitation does not have a noticeable effect on delayed flights in the origin airport.

We find that wind speed has a p value of 0.490046 which is greater than our alpha value of .05 so we would say that precipitation does not have a noticeable effect on delayed flights in the origin airport.


```{r}
modelWeather <- glm(delay ~ temp + dewp + humid + wind_dir + wind_gust + pressure + visib, data = merged)
summary(modelWeather)
```


# Tree-Based Models

## All Factors

### Training

```{r all factors}
set.seed(101)
trainIndex <- sample(1:nrow(merged), 0.7 * nrow(merged))

training_data <- merged[trainIndex, ]
testing_data <- merged[-trainIndex, ]

mod_tree = tree(delay ~ ., data = training_data)
mod_tree

# get model for decision tree
summary(mod_tree)

# plot the tree in a dendrogram
plot(mod_tree, cex = 0.8)
text(mod_tree, pretty = 0, cex = 0.6)

# Find RMSE
mod_tree_pred = predict(mod_tree, newdata = training_data)
mod_tree_rmse = sqrt(mean((training_data$delay - mod_tree_pred)^2))
mod_tree_rmse

# Use K-fold cross validation to reduce tree complexity
cv_tree = cv.tree(mod_tree)
names(cv_tree)

cv_tree

# identify trends in deviance with respect to the size of the tree
with(cv_tree, plot(dev~size, type = 'b'))
```
```{r Random Forest with training}

# fit the random forest model
mod_rf = randomForest(delay ~ ., data = training_data,  mtry = sqrt(NCOL(training_data)-1), importance = TRUE)
mod_rf

plot(mod_rf)

mod_rf_pred = predict(mod_rf, newdata = training_data)
mod_rf_rmse  = sqrt(mean((training_data$delay - mod_rf_pred)^2))
mod_rf_rmse

# run lines together to compare variable importance in bagging vs. RF
varImpPlot(mod_rf, cex = 0.6)

# Compare the three RMSEs to judge which one we will use for testing
cat("Tree RMSE:", mod_tree_rmse, "\n")
cat("Random Forest RMSE:", mod_rf_rmse)


```

### Testing

```{r tree with testing}

set.seed(123)
mod_tree = tree(delay ~ ., data = training_data)
mod_tree

# get model for decision tree
summary(mod_tree)

# plot the tree
plot(mod_tree, cex = 0.8)   # Adjust tree size
text(mod_tree, pretty = 0, cex = 0.6)  # Adjust text size

mod_tree_pred = predict(mod_tree, newdata = testing_data)
mod_tree_rmse = sqrt(mean((testing_data$delay - mod_tree_pred)^2))
mod_tree_rmse

cv_tree = cv.tree(mod_tree)
names(cv_tree)

cv_tree

with(cv_tree, plot(dev~size, type = 'b'))
```

```{r Random Forest with testing}

set.seed(123)
mod_rf = randomForest(delay ~ ., data = training_data,  mtry = sqrt(NCOL(training_data)-1), importance = TRUE)
mod_rf

plot(mod_rf)

mod_rf_pred = predict(mod_rf, newdata = testing_data)
mod_rf_rmse  = sqrt(mean((testing_data$delay - mod_rf_pred)^2))
mod_rf_rmse

varImpPlot(mod_rf, cex = 0.6)

library(rpart)
library(rpart.plot)

tree_model <- rpart(delay ~ ., data = testing_data)
rpart.plot::rpart.plot(tree_model)

# Get the feature importance
importance_matrix <- randomForest::importance(mod_rf)

# Perform hierarchical clustering on feature importance
dendrogram <- hclust(dist(importance_matrix))

# Plot the dendrogram
plot(dendrogram, main = "Dendrogram of Feature Importance", xlab = "Features", ylab = "Distance")

cat("Tree RMSE:", mod_tree_rmse, "\n")
cat("Random Forest RMSE:", mod_rf_rmse)
```

## Month, dep_delay removed

```{r decision tree model}

flights_test <- merged
flights_test <- flights_test %>%
  drop_na()

flights_test <- flights_test[, !names(flights_test) %in% c("month", "dep_delay")]

set.seed(101)
trainIndex <- sample(1:nrow(flights_test), 0.7 * nrow(flights_test))

training_data <- flights_test[trainIndex, ]
testing_data <- flights_test[-trainIndex, ]

mod_tree = tree(delay ~ ., data = training_data)
mod_tree

# get model for decision tree
summary(mod_tree)

# plot the tree in a dendrogram
plot(mod_tree, cex = 0.8)
text(mod_tree, pretty = 0, cex = 0.6)

# Find RMSE
mod_tree_pred = predict(mod_tree, newdata = training_data)
mod_tree_rmse = sqrt(mean((training_data$delay - mod_tree_pred)^2))
mod_tree_rmse

# Use K-fold cross validation to reduce tree complexity
cv_tree = cv.tree(mod_tree)
names(cv_tree)

cv_tree

# identify trends in deviance with respect to the size of the tree
with(cv_tree, plot(dev~size, type = 'b'))



colSums(is.na(testing_data))

nrow(training_data)
nrow(testing_data)

```


```{r forest with training}

set.seed(123)
#colSums(is.na(training_data))

# fit the random forest model
mod_rf = randomForest(delay ~ ., data = training_data,  mtry = sqrt(NCOL(training_data)-1), importance = TRUE)
mod_rf

plot(mod_rf)

mod_rf_pred = predict(mod_rf, newdata = training_data)
mod_rf_rmse  = sqrt(mean((training_data$delay - mod_rf_pred)^2))
mod_rf_rmse

# run lines together to compare variable importance in bagging vs. RF
varImpPlot(mod_rf, cex = 0.6)

# Compare the three RMSEs to judge which one we will use for testing
cat("Tree RMSE:", mod_tree_rmse, "\n")
cat("Random Forest RMSE:", mod_rf_rmse)
```

```{r tree with testing}

set.seed(123)
mod_tree = tree(delay ~ ., data = training_data)
mod_tree

# get model for decision tree
summary(mod_tree)

# plot the tree
plot(mod_tree, cex = 0.8)   # Adjust tree size
text(mod_tree, pretty = 0, cex = 0.6)  # Adjust text size

mod_tree_pred = predict(mod_tree, newdata = testing_data)
mod_tree_rmse = sqrt(mean((testing_data$delay - mod_tree_pred)^2))
mod_tree_rmse

cv_tree = cv.tree(mod_tree)
names(cv_tree)

cv_tree

with(cv_tree, plot(dev~size, type = 'b'))
```


```{r Random Forest with testing}

set.seed(123)
mod_rf = randomForest(delay ~ ., data = training_data,  mtry = sqrt(NCOL(training_data)-1), importance = TRUE)
mod_rf

plot(mod_rf)

mod_rf_pred = predict(mod_rf, newdata = testing_data)
mod_rf_rmse  = sqrt(mean((testing_data$delay - mod_rf_pred)^2))
mod_rf_rmse

varImpPlot(mod_rf, cex = 0.6)

library(rpart)
library(rpart.plot)

tree_model <- rpart(delay ~ ., data = testing_data)
rpart.plot::rpart.plot(tree_model)

# Get the feature importance
importance_matrix <- randomForest::importance(mod_rf)

# Perform hierarchical clustering on feature importance
dendrogram <- hclust(dist(importance_matrix))

# Plot the dendrogram
plot(dendrogram, main = "Dendrogram of Feature Importance", xlab = "Features", ylab = "Distance")

cat("Tree RMSE:", mod_tree_rmse, "\n")
cat("Random Forest RMSE:", mod_rf_rmse)
```





```{r}
carrierDelay <- merged %>%
  filter(dep_delay > 15) %>%
  group_by(carrier) %>%
  summarise(delayNum = n())

ggplot(carrierDelay, aes(x = carrier, y = delayNum, fill = carrier)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Number of Delays by Carrier",
    x = "Carrier",
    y = "Number of Delays"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}
delayPerMonth <- merged %>%
  filter(dep_delay > 15) %>%  
  group_by(month) %>%
  summarize(count = n(), .groups = 'drop')

ggplot(delayPerMonth, aes(x = month, y = count, fill = month)) +
  geom_bar(stat = "identity") +
  labs(title = "Delays by Month",
       x = "Month",
       y = "Number of Delays") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```



```{r}
merged <- merged %>%
  mutate(tempPerDelay = cut(temp,
                        breaks = seq(floor(min(temp)), ceiling(max(temp)), by = 5),
                        include.lowest = TRUE))

delaysTemp <- merged %>%
  filter(dep_delay > 15, !is.na(tempPerDelay)) %>%  
  group_by(tempPerDelay) %>%
  summarize(count = n(), .groups = 'drop')

ggplot(delaysTemp, aes(x = tempPerDelay, y = count, fill = tempPerDelay)) +
  geom_bar(stat = "identity") +
  labs(title = "Delays by Temperature",
       x = "Temperature Range",
       y = "Number of Delays") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
airlineDelay <- merged %>%
  filter(dep_delay > 15) %>%
  group_by(name.x) %>%
  summarise(delayNum = n())

ggplot(airlineDelay, aes(x = name.x, y = delayNum, fill = name.x)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Number of Delays by Airline",
    x = "Airline",
    y = "Number of Delays"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}
merged <- merged %>%
  mutate(depTime = cut(sched_dep_time,
                            breaks = seq(500, 2400, by = 100),
                            include.lowest = TRUE,
                            labels = paste0(seq(500, 2300, by = 100), "-", seq(600, 2400, by = 100))))

delayByTime <- merged %>%
  filter(dep_delay > 15) %>%
  group_by(depTime) %>%
  summarize(count = n(), .groups = 'drop')

ggplot(delayByTime, aes(x = depTime, y = count, fill = depTime)) +
  geom_bar(stat = "identity") +
  labs(title = "Delays by Scheduled Departure Time",
       x = "Scheduled Departure Time",
       y = "Number of Delays") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
merged <- merged %>%
  mutate(windGustperDelay = cut(wind_gust,
                        breaks = seq(floor(min(wind_gust)), ceiling(40), by = 2),
                        include.lowest = TRUE))

delaysTemp <- merged %>%
  filter(dep_delay > 15, !is.na(windGustperDelay)) %>%  
  group_by(windGustperDelay) %>%
  summarize(count = n(), .groups = 'drop')

ggplot(delaysTemp, aes(x = windGustperDelay, y = count, fill = windGustperDelay)) +
  geom_bar(stat = "identity") +
  labs(title = "Delays by Wind Gust",
       x = "Wind Gust",
       y = "Number of Delays") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
